Supporting information: Copulas and their potential for ecology

### 2019-09-29 SupportingInformation.Rmd
part way through edits to results Q1

knitr::kable(tab_3biv_npa_summary, booktabs = T, escape=F,
             format = "latex",linesep = "", row.names=F,
      caption.short = "Nonparametric asymmetry tests against a normal-copula null model with birds, mammals and Cedar Creek data",
      caption = "Nonparametric asymmetry tests against a normal-copula null model with birds, mammals and Cedar creek data: for birds and mammals data, results should be interpreted as of soil C and N results (see Table \\ref{MT-tab_soilCN_asym} which shows stronger upper tail dependence). For Cedar Creek data, first slice ($FS$) and last slice ($LS$) stand for $(0,0.25)$ and $(0.75,1)$ range of lower and upper bounds. Though the non-parametric result does not indicate significantly stronger lower tail dependence for Cedar Creek data, the 
greater than (>) sign before the numbers do show the opposite direction in tail-dependence compared to other bivariate datasets.\\label{tab_3biv_npa_summary}")
```

# Verifying Q1 results for bivariate datasets using nonparametric statistics \label{Q1bivarnonparam}

<!--***DAN: This appendix will eventually be eliminated when we move all the content to the 
main text. The next appendix (S7) will be retained pretty much as is. Probably S8 will all be moved to
the main text and S9 will be kept as it is.-->

The values of the statistics $\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$ and $\Dsq_{l_b,u_b}$ for the soil
C and N data
were compared to distributions of their values on 1000 Kendall- or Spearman-preserving normal surrogates
of the data,
separately in two comparisons for each of the ranges $(l_b,u_b)=(0,0.1),\ldots,(0.9,1)$ 
(Fig. \ref{MT-fig_soilCN_nonparam}). Results show that tail dependence of data is stronger in 
both the lower and upper tails than would be expected under a null hypothesis of a normal copula.
The values of the statistics $\cor_{0,0.1}-\cor_{0.9,1}$, $\Ps_{0,0.1}-\Ps_{0.9,1}$ and
$\Dsq_{0.9,1}-\Dsq_{0,0.1}$ for the soil C and N data were also compared to distributions of their values
on 1000 Kendall- or Spearman-preserving normal surrogates, separately in two 
comparisons (Table \ref{MT-tab_soilCN_asym}). Results showed that
upper-tail dependence was significantly stronger than lower-tail dependence.

For first three bivariate datasets in Table \ref{MT-tab_bivar_summary},
the statistic $P_{0,0.1}-P_{0.9,1}$ was computed for real data and 
for 1000 Kendall-preserving surrogate datasets with normal copulas. Due to relatively smaller sample size,
for Cedar Creek data we used $P_{0,0.25}-P_{0.75,1}$.
For each dataset, the rank of the statistic for real-data in 
the distribution of surrogate values was computed. 
If ranks were $>975$ or $<25$, asymmetry of tail 
dependence was significant compared to a normal-copula null hypothesis. Ranks were `r unname(tab_bivar_summary[10,])`
for, respectively, the soil C and N dataset, the bird body masses and metabolic rates dataset, 
the mammal body masses and metabolic rates dataset, and the Cedar Creek data for the year 2000. 
Thus the first three datasets showed stronger 
upper- than lower-tail dependence, and the Cedar Creek data showed the reverse 
(though not significantly), confirming the 
results based on model selection presented in the main text (see 
Table \ref{MT-tab_bivar_summary}, row 9). Model selection results for Cedar Creek showed 
only slightly stronger lower- than upper-tail dependence, consistent with the non-significant 
tendency toward lower-tail dependence also revealed by the non-parametric analysis for that
dataset.

@@

@@

# Spatial resampling with multivariate datasets for nonparametric statistics \label{spatial_resamp}

>>>>

knitr::kable(tab_3biv_npa_summary, booktabs = T, escape=F,
             format = "latex",linesep = "", row.names=F,
      caption.short = "Nonparametric asymmetry tests against a normal-copula null model with birds, mammals and Cedar Creek data",
      caption = "Nonparametric asymmetry tests against a normal-copula null model with birds, mammals and Cedar Creek data. For bird and mammal data, results should be interpreted in the same way as soil C and N results (Table \\ref{MT-tab_soilCN_asym}). For Cedar Creek data, first slice ($FS$) and last slice ($LS$) stand for the ranges $(0,0.25)$ and $(0.75,1)$ for the lower and upper bounds.\\label{tab_3biv_npa_summary}")
```

@@

<!--***DAN: I'm inclined to eliminate this appendix and put the material
in the main text-->

@@

# Spatial resampling with multivariate datasets for nonparametric statistics \label{spatial_resamp}
<!--***DAN: leave this section exactly as it is, but it may be out of order
now that we describe how this is done in the methods. In fact, I think I
have to do this more generally.-->
